Solution-aware vs global ReLU selection: \\ partial MILP strikes back for DNN verification

### 2025-01-27 CAV 2025 TODO/Comparison.tex
comparison

Then, the scoring functions $\bm{s}_{SR}$ and $\bm{s}_{FSB}$ for ReLUs in layer $k$ are computed by approximating how each node would impact the neurons in layer $n$, by also factoring in the bias $\bm{b}_k$:

\begin{align*}
	\bm{s}_{SR}(k) =& \lvert\max\{0,\bm{\bar{\lambda}}_{k}\odot\bm{b}_{k}\}-\frac{\hat{\bm{u}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}}_k\odot\bm{\bar{\lambda}}_{k}\odot\bm{{b}}_{k}+\frac{\hat{\bm{u}}_k\odot\hat{\bm{l}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}_k}\odot[\bm{\bar{\lambda}}_{k}]_+\rvert \odot\bm{1}_{\hat{\bm{u}}_k>0,\hat{\bm{l}}_k<0} \\
	\bm{s}_{FSB}(k) =& \lvert\min\{0,\bm{\bar{\lambda}}_{k}\odot\bm{b}_{k}\}-\frac{\hat{\bm{u}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}}_k\odot\bm{\bar{\lambda}}_{k}\odot\bm{{b}}_{k}+\frac{\hat{\bm{u}}_k\odot\hat{\bm{l}}_k}{\hat{\bm{u}}_k-\hat{\bm{l}}_k}\odot[\bm{\bar{\lambda}}_{k}]_+\rvert \odot\bm{1}_{\hat{\bm{u}}_k>0,\hat{\bm{l}}_k<0}
\end{align*}

@@

		\node (hidden50) at ($(input1) + (7.5,0.6)$) {$[-1,0.5]$};

@@

		\draw[->,thick] (input1) -- (hidden1) node[near start, above] {$1$};
		\draw[->,thick] (input1) -- (hidden2)node[near start, above] {$-10$};

@@

		\draw[->,thick] (hidden4) -- (hidden5)node[near start, below] {$1$};
		\draw[->,thick] (hidden4) -- (hidden6)node[near start, below] {$1$};

@@

		\draw[->,thick] (output1) -- (output3)node[near start, below] {$-100$};
		\draw[->,thick] (output2) -- (output3)node[near start, below] {$1$};

@@

	% Draw axes
	\draw[->] (-5,0) -- (4,0) node[right] {$x$};
	\draw[->] (0,-1) -- (0,3) node[above] {$y$};

	% Draw ReLU function
	\draw[line width=0.4mm, blue] (-3,0) -- (0,0);
	\draw[thick, blue] (0,0) -- (2.5,2.5) node[below, shift={(0.5,-0.4)}] {$y = \ReLU(x)$};
	\draw[thick, blue] (-3,0) -- (2.5,2.5) node[above, shift={(-5,-1.5)}] {$y = \frac{\UB}{\UB-\LB} x-\frac{\UB\cdot\LB}{\UB-\LB}$};

@@

\end{tikzpicture}
	\caption{}
\label{img:Utility}
\end{figure}

>>>>

Then, the scoring functions $\bm{s}_{SR}$ and $\bm{s}_{FSB}$ for ReLUs in layer $k$ are computed by approximating how each node would impact the neurons in layer $n$, by also factoring in the bias $\bm{b}_k$ in different ways:

\begin{align*}
	\bm{s}_{SR}(k) =& %\bm{1}_{\bm{u}_k>0,\bm{l}_k<0} \odot
	\Biggl\lvert r(\bm{u}_k) \odot \LB(\bm{u}_k) \odot \max(\bm{\lambda}_{k},0)
	+ \max\{0,\bm{\lambda}_{k}\odot\bm{b}_{k}\}-r(\bm{u}_k) \odot\bm{\lambda}_{k}\odot\bm{{b}}_{k}
	\Biggr\rvert  \\
	\bm{s}_{FSB}(k) =& \Biggl\lvert r(\bm{u}_k) \odot \LB(\bm{u}_k) \odot \max(\bm{\lambda}_{k},0)
	+ \min\{0,\bm{\lambda}_{k}\odot\bm{b}_{k}\}-r(\bm{u}_k) \odot\bm{\lambda}_{k}\odot\bm{{b}}_{k}
	\Biggr\rvert
\end{align*}

In case where the bias is 0, then there is no difference between $s_{SR}$ and $s_{FSB}$.
Then, unstable ReLU (with $\LB(u_k) < 0 < \UB(u_k)$) are ranked using these scores, to select the most important ReLUs, as stable ReLUs are linear, thus they are already treated correctly.

Consider the example in Fig. \ref{img:FSB_example}.
We have $\lambda(b)=d$, $\lambda(a)=\frac{cd}{2}$.
Let us do a case analysis:
The most favorable cases are $c < 0 < d$ and $d < 0 < c$: 
as $cd <0$, we have $s_{FSB}(a)=\max(0,\frac{cd}{4}) = 0$.
Now, $a$ and $\hat{a}$ need to be minimized in order to maximize $z$, because $cd <0$. 
For the triangle approximation, having a value $sol(a)$ for $a$, the minimal value for $\hat{a}$ is $\ReLU(sol(a))$. If we open $a$, then the value for $\hat{a}$ will also be $\ReLU(sol(a))$. That is, opening this ReLU is not helping in the approximation of $z$, as correctly predicted by the score.

Case $c>0,d>0$: $s_{FSB}(a)=\frac{cd}{4}$.
Indeed, the value of $a,\hat{a},b,\hat{b}$ should be maximized to maximize the value of $z$, because $W_{bz}=d>0$ and $W_{ab}=c>0$. Hence each variation of $b$ by $\Delta(b)$ will have an impact $r(b) \cdot \Delta(b)$.
Now, let us call $sol(a)$ the maximum value for $a$.
Then the maximum value for $\hat{a}$ is $r(a)\cdot (sol(a)-LB(a))$ according to the triangle approximation when $\ReLU(a)$ is not opened. Now, if $\ReLU(a)$ is opened, then 
the value of $\hat{a}$ is $\ReLU(sol(a))$: the difference is at most $r(a) \cdot LB(a) = \frac{1}{2}$. Going through $b$, the difference is at most $\frac{c}{2}$, which means a difference at most $\frac{c}{4}$ for $\hat(b)$, considering the upper bound of the triangle approximation, meaning a difference at most $\frac{cd}{4}$, which is what is computed by 
$s_{FSB}(a)$. Of course, the difference of opening $\ReLU(a)$ could also be 0 e.g. if $\sol(a)=\LB(a)$ or $\sol(a)=\UB(a)$.

The last case is the most problematic: 
Case $c<0,d<0$, implying that $s_{FSB}(a)=\frac{cd}{4}$.

@@

		\node (hidden50) at ($(input1) + (7.5,0.6)$) {$[-1,1]$};

@@

		\draw[->,thick] (input1) -- (hidden1) node[near start, above] {$c$};
		\draw[->,thick] (input1) -- (hidden2)node[near start, below] {$-10$};

@@

		\draw[->,thick] (hidden4) -- (hidden5)node[near start, above] {$1$};
		\draw[->,thick] (hidden4) -- (hidden6)node[near start, below] {$1$};

@@

		\draw[->,thick] (output1) -- (output3)node[near start, above] {$d$};
		\draw[->,thick] (output2) -- (output3)node[near start, below] {$1$};

@@

		% Draw axes
		\draw[->] (-5,0) -- (4,0) node[right] {$b$};
		\draw[->] (0,-1) -- (0,3) node[above] {$\hat{b}$};

		% Draw ReLU function
		\draw[line width=0.4mm, blue] (-3,0) -- (0,0);
		\draw[thick, blue] (0,0) -- (2.5,2.5) node[below, shift={(-1.1,-1.9)}] {$\hat{b} = \ReLU(b)$};
		\draw[thick, blue] (-3,0) -- (2.5,2.5) node[above, shift={(-6.5,-2)}] {$\hat{b} = \frac{\UB}{\UB-\LB} (b-\LB)$};

	\end{tikzpicture}
		\caption{Triangle approximation}
	\label{img:FSB}
	\end{figure}

	% Draw axes
	\draw[->] (-5,0) -- (4,0) node[right] {$b$};
	\draw[->] (0,-1) -- (0,3) node[above] {$\hat{b}$};

	% Draw ReLU function
	\draw[line width=0.4mm, blue] (-3,0) -- (0,0);
	\draw[thick, blue] (0,0) -- (2.5,2.5) node[below, shift={(-1.1,-1.9)}] {$\hat{b} = \ReLU(b)$};
	\draw[thick, blue] (-3,0) -- (2.5,2.5) node[above, shift={(-6.5,-2)}] {$\hat{b} = \frac{\UB}{\UB-\LB} b-\frac{\UB\cdot\LB}{\UB-\LB}$};

@@

\end{tikzpicture}
	\caption{}
\label{img:Utility2}
\end{figure}
